Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewing English grammar from description of the framework. Couple of typos.

Recent works propose the use of cloud computing services for the components of
client-server distributed EAs. Cloud resources
can reduce operational costs by outsourcing both hardware and software maintenance
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
is available in most laboratories, allowing researches to
scale algorithms at reasonable costs. Researchers have proposed the use of
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
and DropBox~\cite{mericloud}. The use of JavaScript to implement
server side components
gives software developers certain advantages in the cloud
paradigm, since it is a first-class citizen
in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
developers can use the same software tools for both client and server
side development.

@@

In general, a distributed volunteer-based evolutionary computation
framework based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact,
% faster, but the fact that single-user performance is higher (than
% what?) is over the offset. This is because the number of spontaneous
% volunteers we can gather at whim, including people using their
% mobile phones or tablets, is several orders of magnitude higher,
% which is the objective in this kind of systems." (?) 
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

Several kind of threats are possible in this environment; first, since its
programming interface is open and its code is open source, it is
relatively easy to find vulnerabilities and 
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome.
Other two different kinds of attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released
as free software. 
 We also release as an open access paper the result of the 
 experiments performed \cite{DBLP:journals/corr/GuervosG15}  
 and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade it. This
socio-technical approach is, indeed, 
coherent with our whole approach to the study of a distributed volunteer
 system which can only be tackled by approaching it from the
technical as well as the social, as in social media, angle.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the ''CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \cite{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population, then it sends, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request), and requests a random
  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
\end{enumerate}

>>>>

Recent works propose the use of cloud computing services for the components of
client-server distributed EAs. Cloud resources
can reduce operational costs by outsourcing both hardware and software maintenance
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what
is available in most laboratories, allowing researchers to
scale algorithms at reasonable costs. Researchers have proposed the use of
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards},
and DropBox~\cite{mericloud}. The use of JavaScript to implement
server side components
gives software developers certain advantages in the cloud
paradigm, since it is a first-class citizen
in most Platforms as a Service (PaaS) products as the {\tt Node.js} framework is supported by most of them \cite{wood13:nodejs:paas}. Many web services
have native JavaScript libraries too, and also use the JSON (JavaScript Object Notation) format for data interchange. Another advantage is that
developers can use the same software tools for both client and server
side development.

@@

In general, a distributed volunteer-based evolutionary computation
framework based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact,
% faster, but the fact that single-user performance is higher (than
% what?) is over the offset. This is because the number of spontaneous
% volunteers we can gather at whim, including people using their
% mobile phones or tablets, is several orders of magnitude higher,
% which is the objective in this kind of systems." (?) 
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and then compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

Several kind of threats are possible in this environment; first, since its
programming interface is open and its code is open source, it is
relatively easy to find vulnerabilities and 
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome.
Other two different kinds of attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released
as free software. 
 We also released as an open access paper the result of the 
 experiments performed \cite{DBLP:journals/corr/GuervosG15}  
 and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade it. This
socio-technical approach is, indeed, 
coherent with our whole approach to the study of a distributed volunteer
 system which can only be tackled by approaching it from the
technical as well as the social, as in social media, angle.

@@

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the ``CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \cite{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population, then it sends, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request), and requests a random
  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
\end{enumerate}
